### 2015-07-17 ms/ifm.tex
editing

\documentclass[12pt,aps,prb,preprint]{revtex4}
\usepackage{pgf}

\def\ket#1{|#1\rangle}
\def\X{{\bf X}}
\def\A{{\bf A}}

\begin{document}

\title{A Classical Analog of Quantum Interaction-Free Measurement}

\author{Prasenjit Saha}
\email{psaha@physik.uzh.ch}
\affiliation{Physik-Institut, University of Zurich, Switzerland}
\author{Nick Kaiser}
\email{kaiser@hawaii.edu}
\affiliation{Institute for Astronomy, University of Hawaii at Manoa,
Honolulu, HI 96822}

\date{\today}

\begin{abstract}
Interaction-free measurement generally refers to quantum phenomena
such as the presence of a photon absorber being inferred without it
actually absorbing any photons.  Here we make the simple observation
that there is an analogous situation in classical mechanics: two
coupled oscillators, one of which is damped; the energy dissipation
can be made arbitrarily small by making the damping large enough, thus
revealing the presence of the damper despite arbitrarily low energy
dissipation.
\end{abstract}

\maketitle

\section{Introduction}

Many quantum paradoxes have been explored theoretically and
experimentally in recent years.  Most involve entangled states: the
famous EPR paradox is an old example, and quantum computing is a more
recent development\cite{2007qucosc.book}.  Entanglement is considered
an intrinsically quantum phenonemon with no classical analog.

There are also counter-intuitive quantum phenonema not involving
entanglement.  Imagine a photon, initially in a ``vertical'' polarized
state.  It is sent through a Faraday rotator, and then through a
vertically polarizing sheet.  The former rotates the polarization
plane towards ``horizontal'' by a small angle, say $\pi/(2N)$, where
$N$ is large.  The latter effectively measures the polarization state
in the vertical/horizontal basis; if the result comes out horizontal,
the photon is absorbed, otherwise it goes through in a vertical
eigenstate.  Since $N$ is large, the chances of being absorbed are
small.  More specifically, after the rotation the {\em amplitude\/} to
be in the horizontal state will be $\sim1/N$ so the probability to be
in that mode --- and therefore the probability of absorption --- will
be $\sim1/N^2$.  Thus, most likely, the photon will get slightly
rotated and then reset to vertical.  If the photon is re-routed back
to the apparatus $N$ times, it will still most likely emerge
vertically polarized.  The rotator and the polarizer appear to cancel
each other out.  The general phenomenon is known as the quantum Zeno
effect,\cite{1977JMP....18..756M,PhysRevA.41.2295} but it can also be
described in everyday language, in terms of waking and sleeping
states.\cite{orzel2009teach}

What happens if we now remove the polarizer?  Clearly, the $N$ Faraday
rotations will then convert the photon to a horizontal state.  Thus,
the net effect of the polarizer is to rotate a photon from horizontal
to vertical.  Despite the fact that the polarizer is really just an
absorber of horizontal photons, it can, within this experimental
set-up, have the effect of rotating photons instead of absorbing them.
This is the essence of interaction-free measurement; we can infer, or
`measure' the presence or absence of the polarizer by measuring the
polarization.  If the polarization is rotated we know that the
absorber cannot have been present -- and there is then of course no
absorption -- while if the polarization is not changed then we know
that the absorber was present, but again, at least in the limit $N
\rightarrow \infty$, no absorption actually takes place.

Since the initial suggestion (essentially the $N=1$
case)\cite{1993FoPh...23..987E} and generalization to the now-standard
form, together with the first experiments\cite{PhysRevLett.74.4763}
interaction-free measurement has been the subject of many
demonstrations\cite{PhysRevA.57.3987,voorthuysen:1504,2006JPhB...39.3177N}
and reviews.\cite{deweerd:272} (An older usage of similar
terminology\cite{springerlink:10.1007/BF01327019,dicke:925} refers to
a situation where a non-detection enables an inference.)  Applications
to non-invasive medical imaging\cite{PhysRevA.80.040902} are
particularly interesting.

In this paper, we argue that interaction-free measurement need not be
quantum at all.  Rather, the characteristics features of (a)~an
amplitude that cycles between two states, but (b)~an absorber can
block this cycling, thus revealing its presence, but without actually
absorbing any energy, can also appear in classical systems.  If one
has a pair of identical coupled oscillators $x$ and $y$, and $x$ is
set in motion, one will see the strength of the oscillation of $x$
slowly decrease with time, and over longer times the energy will
periodically slosh back and forth between the two systems.  But if
instead oscillator $y$ is damped --- so it is able to {\em absorb\/}
energy --- the rate of decrease of the energy in oscillator $x$ will
be suppressed.  Thus one can infer the ability of oscillator $y$ to
absorb energy by observing the behaviour of oscillator $x$ without, in
the limit that the damping is strong, there actually being any energy
absorbed.

In the other sections of this paper, we flesh out the above argument.
In section~\ref{sec:interf} we work through an alternative well-known
description of interaction-free measurement, more closely resembles
some of the actual experiments.  This is in terms of a Mach-Zehnder
interferometer, whose two arms replace the horizontal and vertical
polarization states of the above thought-experiment.
Figure~\ref{fig:mz} schematically shows the $N=1$ case; the general
case would be a concatenation of such devices.  With an absorber, the
amplitude sloshes between the 'lower' and 'upper' arms of the
interferometer; the absorber results in confining the amplitude to the
lower arm.  Section~\ref{sec:oscill} briefly solves for the
oscillators $x$ and $y$ from the previous paragraph.

\section{An interferometer as an absorber detector}\label{sec:interf}

Let us express the interferometric setup of Figure~\ref{fig:mz} in the
language of quantum two-state systems, borrowing the notation of
quantum computing.  Figure~\ref{fig:qubits} replaces the optical
elements with states and operators.  The two eigenstates are rays
moving towards upper right, or towards lower right in the figures. We
write these states as kets or matrices as
\begin{equation}
\ket0 \equiv \left( \begin{array}{c} 1 \\ 0 \end{array} \right)  \qquad
\ket1 \equiv \left( \begin{array}{c} 0 \\ 1 \end{array} \right)
\end{equation}
while
\begin{equation}
\X \equiv \left( \begin{array}{cc} 0 & 1 \\ 1 & 0 \end{array} \right)
\end{equation}
denotes an operator swapping these states.  A beam splitter is given
by the unitary operator
\begin{equation}
\exp\left[i\left({\textstyle\frac\pi2-\epsilon}\right)\right] \X \equiv
\sin\epsilon + i\cos\epsilon\,\X
\end{equation}
Then $\cos^2\epsilon$ is the reflectivity.  Small $\epsilon$ means
highly reflective, and $\epsilon=0$ corresponds to a mirror, whose
operator is $i\X$.  We emphasize, however, that the quantum notation
does not mean the interferometer is intrinsically quantum.  The
behavior of such an interferometer can also be understood in terms of
classical waves obeying Maxwell's equations.

Consecutive operations of beam-splitting, reflection and beam
splitting can be written as
\begin{equation}
     (\sin\epsilon + i\cos\epsilon\,\X)
     \; i\X \;
     (\sin\epsilon + i\cos\epsilon\,\X)
\end{equation}
which simplifies to
\begin{equation}
     -(\sin2\epsilon + i\cos2\epsilon\,\X)
\label{eqn:no-absorber}
\end{equation}
and is illustrated by the upper and middle diagrams in
Figure~\ref{fig:qubits}.

If we now add more elements to the interferometer, specifically $N-1$
sets of beam-splitters alternating with mirrors, the resulting state
is
\begin{equation}
     \left[ (\sin\epsilon + i\cos\epsilon\,\X) \; i\X \right]^{N-1} \;
     (\sin\epsilon + i\cos\epsilon\,\X)
\end{equation}
which simplifies to
\begin{equation}
     (-1)^{N-1} (\sin N\epsilon + i\cos N\epsilon\,\X)  \,.
\end{equation}
If $N\epsilon=\pi/2$ this is identity, apart from a phase factor.

Now let us put an absorber in one of the arms.  This blocks the
$\ket0$ state, while the mirror (in the other arm) reflects $\ket1$ as
before.  Then we will have
\begin{equation}
     \left[ (\sin\epsilon + i\cos\epsilon\,\X) \; i\A \right]^{N-1} \;
     (\sin\epsilon + i\cos\epsilon\,\X)
  \qquad
   \A \equiv \left( \begin{array}{cc} 0 & 1 \\ 0 & 0 \end{array} \right)
\end{equation}
involving a non-Hermitian operator $\A$.  Working out the algebra
gives
\begin{equation}
(-1)^{N-1} \, \cos^N\!\epsilon \,
\left( \begin{array}{cc} \tan\epsilon & -i\tan^2\epsilon \\
                         i            & \tan\epsilon 
       \end{array} \right)
\label{eqn:absorber}
\end{equation}
If $\epsilon$ is small (very slightly transmissive) this operator
changes $\ket0$ to $\ket1$.  The lower diagram in
Figure~\ref{fig:qubits} illustrates the result for $N=1$.

The contrasting results (\ref{eqn:no-absorber}) and
(\ref{eqn:absorber}) express the characteristic feature of
interaction-free measurement.  With no absorber the amplitude cycles
between $\ket0$ and $\ket1$.  An absorber blocks the sloshing, without
necessarily absorbing.

\section{Coupled oscillators as an absorber detector}\label{sec:oscill}

In the previous section we considered a system that could be quantum,
or classically wave-optic.  In this section, we show that the same
qualitative behavior emerges in a purely classical system.

Consider two coupled oscillators, one of them damped.  Such a system
can be written, with a suitable choice of units, as
\begin{equation}
\begin{array}{l}
\ddot x + x + \kappa y = 0 \\
\ddot y + y + \kappa x + \gamma y = 0 \,.
\label{eqn:coupled}
\end{array}
\end{equation}
Without the damping term ($\gamma=0$), it is easy to verify that
\begin{equation}
\begin{array}{l}
2x = \cos(\sqrt{1+\kappa}\,t) + \cos(\sqrt{1-\kappa}\,t)  \\
2y = \cos(\sqrt{1+\kappa}\,t) - \cos(\sqrt{1-\kappa}\,t)
\end{array}
\end{equation}
is a solution for the initial conditions
\begin{equation}
x(0) = 1, y(0) = \dot x(0) = \dot y(0) = 0 \,.
\label{eqn:inic}
\end{equation}
If the coupling is weak ($\kappa\ll1$) the solution takes the form
\begin{equation}
\begin{array}{l}
x = \cos\left(\frac12\kappa t\right) \cos t \\
y = -\sin\left(\frac12\kappa t\right) \sin t \\
\end{array}
\end{equation}
and the amplitude sloshes from $x$ to $y$.

Now putting back the damping, one can solve the equations in different
ways.  A concise way is to use Laplace transforms.  Taking the Laplace
transform of the equations (\ref{eqn:coupled}) with the same initial
conditions (\ref{eqn:inic}) gives
\begin{equation}
\begin{array}{l}
s^2 X - s + X + \kappa Y = 0 \\
s^2 Y + Y + \kappa X + \gamma s Y = 0 \,.
\end{array}
\end{equation}
For strong damping, we have to $O(1/\gamma)$
\begin{equation}
\begin{array}{l}
\displaystyle
X = \frac s{s^2+1} + \frac{\kappa^2}\gamma \frac1{(s^2+1)^2} \\
\displaystyle
Y = -\frac\kappa\gamma \frac1{s^2+1}
\end{array}
\end{equation}
Taking the inverse Laplace transform and simplifying gives the
solution to $O(1/\gamma)$
\begin{equation}
\begin{array}{l}
\displaystyle
x = \cos t + \frac{\kappa^2}{2\gamma} (\sin t - t\cos t) \\
\displaystyle
y = -\frac\kappa\gamma \sin t \,.
\end{array}
\end{equation}
As we see, damping slows down the transfer to $y$.

In a mechanical system, it may seem odd that more damping means less
energy dissipation.  But for electrical oscillotors (an LC circuit
coupled to an RLC circuit) we are used to the idea that increasing
resistance reduces dissipation.  Quantum interaction-free measurement
can thus be seen simply as a quantized version of such a system.

\bibliography{ifm.bib}

\newpage

\def\vmargin{\hrule height 1cm width 0pt}

\def\pgfbeamsplitter(#1,#2)#3{\pgfputat{\pgfxy(#1,#2)}
                              {\pgfbox[center,center]#3
                               \pgfxyline(-1.1,0)(0,1.1)
                               \pgfxyline(0,1.1)(1.1,0)
                               \pgfxyline(1.1,0)(0,-1.1)
                               \pgfxyline(0,-1.1)(-1.1,0)}}

\def\pgfbeamsplitter(#1,#2)#3{\pgfputat{\pgfxy(#1,#2)}
                              {\pgfbox[center,center]#3
                               \pgfxyline(-1.1,0)(0,1.1)
                               \pgfxyline(0,1.1)(1.1,0)
                               \pgfxyline(1.1,0)(0,-1.1)
                               \pgfxyline(0,-1.1)(-1.1,0)}}

\def\semireflecting{$\hbox{beam}\atop\hbox{splitter}$}
\def\arbireflecting{$e^{i\alpha\X}$}
\def\nearlytransparent{$e^{i\left(\frac\pi2-\epsilon\right)\X}$}

\def\pgfsimpmirror(#1,#2){\pgfputat{\pgfxy(#1,#2)}
                      {\pgfbox[center,center]{M}
                      \pgfxyline(-.4,-.4)(.4,-.4)
                      \pgfxyline(.4,-.4)(.4,.4)
                      \pgfxyline(.4,.4)(-.4,.4)
                      \pgfxyline(-.4,.4)(-.4,-.4)}}

\def\pgfmirror(#1,#2){\pgfputat{\pgfxy(#1,#2)}
                      {\pgfbox[center,center]{$i\X$}
                      \pgfxyline(-.4,-.4)(.4,-.4)
                      \pgfxyline(.4,-.4)(.4,.4)
                      \pgfxyline(.4,.4)(-.4,.4)
                      \pgfxyline(-.4,.4)(-.4,-.4)}}
\def\pgfabsorber(#1,#2){\pgfputat{\pgfxy(#1,#2)}
                      {\pgfbox[center,center]{$\bf 0$}
                      \pgfxyline(-.4,-.4)(.4,-.4)
                      \pgfxyline(.4,-.4)(.4,.4)
                      \pgfxyline(.4,.4)(-.4,.4)
                      \pgfxyline(-.4,.4)(-.4,-.4)}}

\begin{figure}
\begin{center}
\vmargin
\begin{pgfpicture}
\pgfbeamsplitter(0,0)\semireflecting
\pgfsimpmirror(2,2) 
\pgfsimpmirror(2,-2)
\pgfbeamsplitter(4,0)\semireflecting
\pgfsetendarrow{\pgfarrowsingle}
\pgfxyline(-1.5,-1.5)(-.7,-.7) 
\pgfxyline(.7,.7)(1.5,1.5)
\pgfxyline(.7,-.7)(1.5,-1.5)
\pgfxyline(2.5,1.5)(3.3,.7) 
\pgfxyline(2.5,-1.5)(3.3,-.7)
\pgfxyline(4.7,.7)(5.5,1.5) 
\end{pgfpicture}
\vmargin\vmargin\vmargin
\begin{pgfpicture}
\pgfbeamsplitter(0,0)\semireflecting
\pgfsimpmirror(2,2) 
\pgfsimpmirror(2,-2)
\pgfbeamsplitter(4,0)\semireflecting
\pgfsetendarrow{\pgfarrowsingle}
\pgfxyline(-1.5, 1.5)(-.7, .7) 
\pgfxyline(.7,.7)(1.5,1.5)
\pgfxyline(.7,-.7)(1.5,-1.5)
\pgfxyline(2.5,1.5)(3.3,.7) 
\pgfxyline(2.5,-1.5)(3.3,-.7)
\pgfxyline(4.7,-.7)(5.5,-1.5) 
\end{pgfpicture}
\vmargin\vmargin\vmargin
\begin{pgfpicture}
\pgfbeamsplitter(0,0)\semireflecting
\pgfabsorber(2,2) 
\pgfsimpmirror(2,-2)
\pgfbeamsplitter(4,0)\semireflecting
\pgfsetendarrow{\pgfarrowsingle}
\pgfxyline(-1.5,-1.5)(-.7,-.7) 
\pgfxyline(.7,.7)(1.5,1.5)
\pgfxyline(.7,-.7)(1.5,-1.5)
\pgfxyline(2.5,-1.5)(3.3,-.7)
\pgfxyline(4.7,.7)(5.5,1.5) 
\pgfxyline(4.7,-.7)(5.5,-1.5) 
\end{pgfpicture}
\end{center}
\caption{\label{fig:mz} Schematic representation of interaction-free
  measurement. Each diagram stands for a Mach-Zehnder set up so that
  both arms are exactly equal.  In the upper and middle diagrams, a
  photon sent axially into one arm emerges parallel to the original
  direction.  In the lower diagram, one of the mirrors is replaced by
  an absorber, and as a result, photons can emerge perpendicular to
  their original direction.  Such photons effectively detect the
  absorber without getting absorbed.}
\end{figure}

\begin{figure}
\begin{center}
\vmargin
\begin{pgfpicture}
\pgfbeamsplitter(0,0)\nearlytransparent 
\pgfmirror(2,2) 
\pgfmirror(2,-2)
\pgfbeamsplitter(4,0)\nearlytransparent 
\pgfsetendarrow{\pgfarrowsingle}
\pgfxyline(-1.5,-1.5)(-.7,-.7) 
\pgfxyline(.7,.7)(1.5,1.5)
\pgfxyline(.7,-.7)(1.5,-1.5)
\pgfxyline(2.5,1.5)(3.3,.7) 
\pgfxyline(2.5,-1.5)(3.3,-.7)
\pgfxyline(4.7,.7)(5.5,1.5)
\pgfxyline(4.7,-.7)(5.5,-1.5)
\pgfputat{\pgfxy(-1.7,-1.7)}{\pgfbox[right,center]{$\ket0$}} 
\pgfputat{\pgfxy(5.7, 1.7)}{\pgfbox[left,center]
         {$-\sin2\epsilon\;\ket0$}}
         \pgfputat{\pgfxy(5.7,-1.7)}{\pgfbox[left,center]
         {$-i\cos2\epsilon\;\ket1$}}
\end{pgfpicture}
\vmargin\vmargin\vmargin
\begin{pgfpicture}
\pgfbeamsplitter(0,0)\nearlytransparent 
\pgfmirror(2,2) 
\pgfmirror(2,-2)
\pgfbeamsplitter(4,0)\nearlytransparent 
\pgfsetendarrow{\pgfarrowsingle}
\pgfxyline(-1.5, 1.5)(-.7, .7) 
\pgfxyline(.7,.7)(1.5,1.5)
\pgfxyline(.7,-.7)(1.5,-1.5)
\pgfxyline(2.5,1.5)(3.3,.7) 
\pgfxyline(2.5,-1.5)(3.3,-.7)
\pgfxyline(4.7,.7)(5.5,1.5)
\pgfxyline(4.7,-.7)(5.5,-1.5)
\pgfputat{\pgfxy(-1.7, 1.7)}{\pgfbox[right,center]{$\ket1$}} 
\pgfputat{\pgfxy(5.7, 1.7)}{\pgfbox[left,center]
         {$-i\cos2\epsilon\;\ket0$}}
          \pgfputat{\pgfxy(5.7,-1.7)}{\pgfbox[left,center]
         {$-\sin2\epsilon\;\ket1$}}
\end{pgfpicture}
\vmargin\vmargin\vmargin
\begin{pgfpicture}
\pgfbeamsplitter(0,0)\nearlytransparent 
\pgfabsorber(2,2) 
\pgfmirror(2,-2)
\pgfbeamsplitter(4,0)\nearlytransparent 
\pgfsetendarrow{\pgfarrowsingle}
\pgfxyline(-1.5,-1.5)(-.7,-.7) 
\pgfxyline(.7,.7)(1.5,1.5)
\pgfxyline(.7,-.7)(1.5,-1.5)
\pgfxyline(2.5,-1.5)(3.3,-.7)
\pgfxyline(4.7,.7)(5.5,1.5)
\pgfxyline(4.7,-.7)(5.5,-1.5)
\pgfputat{\pgfxy(-1.7,-1.7)}{\pgfbox[right,center]{$\ket0$}} 
\pgfputat{\pgfxy(5.7, 1.7)}{\pgfbox[left,center]
         {$-\sin\epsilon\cos\epsilon\;\ket0$}}
          \pgfputat{\pgfxy(5.7,-1.7)}{\pgfbox[left,center]
         {$-i\cos^2\epsilon\;\ket1$}}
\end{pgfpicture}
\end{center}
\caption{\label{fig:qubits} Like Figure \ref{fig:mz} but annotated
  with states and operators.  A photon is denoted by $\ket0$ if moving
  towards upper right, and by $\ket1$ if moving towards lower right.
  Beam splitters and mirrors are unitary operators.  An absorber is a
  non-unitary operator, which zeros the $\ket0$ subspace.}
\end{figure}

\end{document}